Experiences Gamifying Developer Practices

### 2013-10-11 paper_id1.tex
Data Complete

\title{Experiences Gamifying Code Navigation}
\subtitle{Are development practices a habit?}
% 
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
As software development practices evolve, toolsmiths face the continuous challenge of communicating new practices and tools to the community.  Developers both like to try new things and like to stick with the familiar.  In addition to training, discussing, and presenting software engineering practices and tools, we propose a method to both motivate and monitor their adoption through a tool embedded in the Integrated Development Environment (IDE).

Blaze monitors developer practice and tool use over the long term and reports the lasting change a communication effort makes in everyday habits.  Embeding Blaze in the developer`s work environment as a Visual Studio extension captures developer work patterns while giving feedback to users on their practices.   Blaze encourages everyone to use the new practices by assigning points to specific commands and tools.  
 With a rich data source, we evaluate developer practices, define metrics that give individual feedback, and create an environment for developers to see how they are doing compared with their teammates.

Data collected from the pilot group show a dichotomy of navigation and code understanding practices used by developers even the same developer in different situations.  We observed that the duration of an edit session is more correlated with categories of activity that benefit from improved practices and tools. The Blaze tool was well received by most participants, however, the intended improvements in navigation practices promoted by the pilot did not occur.  We conclude that more proactive interventions are necessary to move developers to change their code navigation habits.
\end{abstract}

@@

\section{Introduction}
%update

%As  a developer I'm focused on the problem I'm currently solving and driven by productivity and deadlines to the point that I don't have time to sharpen the saw identifying better practices.

Developers with years of experience develop habits and ways of doing things that are difficult to change even when they might recognize better methods.   For instance, the commands they use and ways they use to navigate the code base in the development environment may make them less efficient than using newer practices. Robillard et. al. show that developers who use commands that allow them to follow code structure are more effective and efficient than developers who browse through code \cite{wbsnipes:Robillard2004How}.  Can we make developers more productivity by training them to use structured navigation commands and tools?

Perhaps the persistence of their habit is why developers are hard to turn from their old practices.  Developers may need some coaching to continue a new practice following an introduction.  To continuously coach the developer, we conceived of a tool in our prior work to provide constant feedback and motivate them to adopt a new practice.  \cite{Snipes2013Towards} This tool creates a competitive game that gives continous feedback to the developer continuously on adopting the improved practice.  The more they change their habits, the higher they score in the game. 
%clunky
Taking a step back, gamification is a popular movement that people apply to customer oriented web sites and social games.  Why would software developers writing serious software be interested in this?  Our first step was to challenge this question through a survey that queried both whether developers would be interested in gamifying their workspace and what aspects of that space would be most interesting to them.

@@

The rest of the paper is organized as follows:  Section 2 reviews the preliminary survey-based assement, Section 3 Discusses the design considerations for developing the Blaze tool/game, Section 4 Discusses the design of the study and participant selection, Section 5 is Related work, Section 6 results from the pilot, Section 7 Threats to Validity, and Section 8 has our Conclusions.

@@

Developers were asked to rank awards for different actions by which they would find the most desirable.  Overall 68\% of developers ranked receiving an award for consistently performing quality practices as either 1st or 2nd place.  This supports a conclusion that focusing on awarding quality practices would motivate developers to do them consistently.
%EMH connection?
\begin{itemize}
\item 95\% of developers responding would try tools and practices suggested by an automated recommendation system particularly if they were provided in context with the current workflow pattern the developer follows.
\item Responses indicate a preference for team competition over individual awards (badges) and awards for quality practices ranked highest.  So, the tool will incorporate team competition ideas and recognize quality practices.
\item Over 90\% of respondents approved of sharing data with their team members
\end{itemize}

@@

@@

Using these guidelines we assigned points to structured navigation commands which received 1-10 points each while unstructured navigation commands received 0 points.    In creating levels, following the guidance from Hamari \cite{wbsnipes:Hamari2011Framework}, we built levels following an exponential curve.  Users initially level up after a working day or two extending to achieving the next level after a week or even a month following an exponential curve.  The levels make the game motivating for beginners and challenging for more experienced developers.  We designed the points and levels such that above average developers can complete all levels during the study period.

@@

After a week's data collection marking the end of the first stage, Blaze started to automatically pop-up a window containing a button for a web page with information.  The page included a link to download another tool for code search called Sando \cite{Shepherd2012Sando} and information about using that tool in a video.  The sub-page on navigation contained a page of tutorial on the built-in structured navigation commands in Visual Studio.  The About Blaze page contained details of the point scoring system and background on Blaze's purpose.  The usefulness of the communication site is rated as part of our study completion survey.

@@

\subsubsection{Observations About Developers` Habits}

We are interested in understanding whether developers are consistently using practices that make them inefficient.  One way to check that is to determine whether our measurements are associated with specific developers in the study, in other words, can we independently classify the edits performed by each developer.  If this is possible, then we can discuss whether developer have consistent ways of navigating and understanding for each edit.  Thus the hypothesis is developers have consistent ways of navigating and understanding code while they are making changes.

To test this hypothesis we construct a classification tree using Weka \cite{Hall2009WEKA} with the J48 classifier.  Slicing the data and training the model using 10-fold cross-validation, the model identifies 53\% of the 647 edit sessions with the correct developer among 6 participants.  The tree uses nearly all available fields including time for all categories and time of day for edits in its attempt to associate sessions with their developer generating a Kappa statistic of 0.403  and square root of the relative squared error at 101\%.

While there is some consistency to how developers work, we cannot say that individual developers are always consistent in their work patterns.  Thus developers are somewhat flexible in the commands they use when navigating and making changes to code and may be more open to adopting new tools and techniques.\

@@

Using Weka\cite{Hall2009WEKA}, we performed an attribute selection process to identify the most significant attributes related to edit session duration.  The attribute search method we chose was Greedy Step-wise search that selects attributes based on their correlation with the target and stops adding attributes when the next attribute decreases the correlation.    The process identified the top 5 out of 10 possible categories as follows:
\begin{enumerate}
\item Unstructured Navigation
\item Read
\item Build
\item Debug
\item Edit
\end{enumerate}

Using R\cite{Rcitation}, we constructed a linear regression model for Unstructured Navigation to determine the correlation with edit duration.  The linear model found a positive correlation with unstructured navigation with a p-value <0.001 on 646 degrees of freedom.  The model with only unstructured navigation had an R-squared of 0.52 showing a good portion of the variation is explained by that category.  In order to have a more acceptable explanation of edit duration, we conducted a multiple linear regression model with both Unstructured Navigation and Read categories. The Read category of events captures the activities the developer is doing to look through the code in the editor.  They are navigation commands that move incrementally like pressing the up/down arrow keys, or moving the scroll-bar an incremental amount.  Combining these two categories results in a linear model with a p-value <0.001 and an R-squared of 0.76, more acceptable for drawing conclusions.

Conducting an ANOVA analysis shown in Table \ref{editDurationANOVA} for this regression model, we can explore the contribution of Unstructured Navigation and Reading to the model.  The contribution of each variable can be determined by calculating the portion of the total variation that each variable explains.  Unstructured Navigation explains 53\% of the variation.  Out of the remaining variation found by subtracting the amount explained by Unstructured Navigation from the total, Read explains 49\%.

\begin{table}[!t]

\begin{mdframed}[linecolor=white]
\begin{tabular}{|l|r|r|r|r|r|}
\hline
                          &  Df  & SSQ & MeanSq &Fval.   & Pr(>F)    \\
\hline
Unstr. Nav.& 1 & 0.039& 0.03920 & 1383 & <0.001\\
\hline
Read                 & 1 & 0.017 & 0.01741 & 614 & <0.001\\
\hline
Residuals                & 645 & 0.018 & 0.00003   &  &  \\
\hline
Total		& 647 & 0.074 & 0.05664 & & \\
\hline
\end{tabular}
\caption{ANOVA for Edit Duration}
\label{editDurationANOVA}
\end{mdframed}\end{table}

After this analysis we cannot disprove that unstructured navigation is correlated with edit duration, thus it is an important factor in developer productivity.  Combining unstructured navigation with reading describes a larger portion of the overall edit duration and motivates us towards improvements in both navigation methods and code comprehension.

\subsubsection{Observations About navigation ratio}

@@

Structured navigation events included: 
\begin{itemize}
\item Navigate To (Ctrl+,) is a fuzzy search interface that locates symbols matching your search terms.
\item Go To Definition (F12) brings up the code where the selected symbol is defined. 
\item View Call Hierarchy (Ctrl+K Ctrl+T) provides a two way analysis of a symbol's dependencies and uses. 
\item Class View (Ctrl+W, C) priovides a searchable browser of classes and class hierarchy
\item Find All References provides a list of refernces for an object
\item Navigate to Event Handler brings up the event handler from the XAML editor
\item View Class Diagram generates a class diagram for the project
\item View Obeject Browser is a search tool for objects in the solution
\end{itemize}

@@

During the pilot we staged the deployment of information feedback to developers so they would gradually improve their structured navigation practices.  Our hypothesis was that this instant feedback and information would result in an increase in the use of structured navigation commands in Visual Studio.

To evaluate whether developers use more structured navigation when they receive instant feedback, we monitor the change in navigation ratio over the pilot study period.   Expecting that as developers learn about navigation commands and tools that count towards the score in the second week they will begin using the practices more.  Also expecting that noticing the leaderboard competition will encourage some developers to use the navigation practices more.

\begin{figure}
	\includegraphics[width=3.3in]{navratioboxplot.pdf}
	\caption{Navigation Ratio by Week}
	\label{fig:navigationaverage}
\end{figure}

What we observed visually in Figure \ref{fig:navigationaverage} is that the structured navigation practice did not increase significantly from the beginning to the end of the study.  Overall, we did not see a significant effect towards increasing navigation ratio for the interventions in week 2 or week 3.   Weeks 4 and 5 that follow the formal period of study show the navigation ratio continues in a similar distribution pattern.

@@

Anil this is one place  you can discuss the findings from the post-interviews with the feedback.  It would be great if you can provide some quotes or anecdotes here.

@@

Following a team oriented analysis of data, we conducted one on one feedback sessions with the pilot participants.  The participants received the feedback in the form of charts and written analysis of the charts display.  The key analysis chart compared their distribution of time in each category to the median for the group.  Figure \ref{fig:developercomparison} shows the chart analyzed for one developer comparing their time to the median time.

We discovered during these interviews that developers did not spend the entire time in their desktop environment because they had a virtual machine environment to use for testing.  Thus categories like testing, debug, and reading were affected by whether or not they used the virtual environment or their desktop for these activities.  Individual differences in those categories are thus not relevant for the feedback given. Because our points analysis focused on navigation commands excluding reading, the data from the alternative environment does not impact the results.  None of the developers said they used Sando in the alternative environment, thus results for points driven by Sando usage remain relevant.

Defelopers also reiterated that they wanted different feedback from the tool and experience such as immediate prompts to suggest navigation commands within the context of what they are working on.  Though commonly held opinion of Microsoft`s Clippy experience suggests this more invasive form of feedback is not always welcome, we nonetheless shall list as future work providing more immeidate suggestions on how developers can improve their navigation practices.

Moving towards more immediate feedback, we shared a set of achivement badges tied to specific navigation commands  with the developers.  Developers thought the bades were interesting and helpful in their specificity.

@@

Our post pilot survey attempts to shed some light on the reasons behind the consistency in navigation ratio over the pilot.

On the assessment of structured navigation knowledge, half of the participants replied that they had very good prior knowledge about structured navigation and they use structured navigation as much as possible.  Other half of the participants have some understanding on structured navigation use it some of the time. Even developers whose navigation ratio  was lower than average, used structured navigation occasionally as much as their more frequently using colleagues.  While analyzing the points they scored with Blaze tool, we observed that the participants who mentioned that they have good knowledge in structured navigation are on top of the points table. 
%the following statement is not supported with data unless you have additional data?
Incidentally, while cross verifying their performance on the coding, they are better productive compared to the participants on the bottom of points table.
\begin{figure}
	\includegraphics[width=3.4in]{ElementInfluenceChart.pdf}
	\caption{Influence of Feedback Elements}
	\label{fig:elementInfluence}
\end{figure}

@@

%Though not significant, there is a loose correlation between those who rated their prior knowledge of structured navigation commands highly and their point scores.   
There was also feedback that people wanted more information presented to them by the tool on what structured navigation commands to use, so they may not have closed the knowledge gap during the pilot.

>>>>

\title{Experiences Gamifying Developer Practices}
%\subtitle{}
% 
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
As software development practices evolve, toolsmiths face the continuous challenge of communicating new practices and tools to the community.  In addition to training, discussing, and presenting software engineering practices and tools, we propose a method to both motivate and monitor their adoption through a tool embedded in the Integrated Development Environment (IDE).

Blaze monitors developer practice and tool use over the long term and reports the lasting change a communication effort makes in everyday habits.  Embedding Blaze in the developer`s work environment as a Visual Studio extension captures developer work patterns while giving feedback to users on their practices.   Blaze encourages everyone to use the new practices by assigning points to specific commands and tools.  
 With a rich data source, we evaluate developer practices, define metrics that give individual feedback, and create an environment for developers to see how they are doing compared with their teammates.

Data collected from the pilot group show a dichotomy of navigation and code understanding practices used by developers even the same developer in different situations.  We observed that the duration of an edit session is more correlated with categories of activity that benefit from improved practices and tools. The Blaze tool was well received by most participants, however, navigation practices were not significantly affected by the Blaze feedback mechanism.  We conclude that more proactive interventions are necessary to encourage developers to change their code navigation habits.
\end{abstract}

@@

\section{Introduction}

Software developers who maintain code face additional challenges of navigating an existing code base to incorporate changes as they build new features or fix bugs.  Over the years, developers learn techniques to analyze code and locate areas that are relevant to an enhancment.  As new techniques and tools are introduced to the development environment, developer may not learn about them unless they interact with a peer developer who uses them\cite{wbsnipes:Hill2011Peer}.    For instance, the commands they use to navigate the code base in the development environment may make them less efficient than using newer practices or tools.  Robillard et. al. show that developers who use commands that allow them to follow code structure are more effective and efficient than developers who browse through code \cite{wbsnipes:Robillard2004How}.

Enter the toolsmith,  who responds to challenges faced by software developers by creating tools such as code search tools that make the job easier and more efficient.  The toolsmith needs to spread the word about advanced tools they develop and determine their effect on developer practices over the long term.  Is there something we can do that encourages developers to adopt new practices and tools that also provides feedback to developers and toolsmiths on improvements that result?

We envisioned a tool in our prior work that would motivate developers to adopt new practices or tools by providing constant feedback and allowing them to compare their improvements to how their peers are doing  \cite{Snipes2013Towards}.  This tool creates a competitive game that gives continuous feedback to the developer on adopting the improved practice.  The more they change their practices to adopt new techniques or tools, the higher they score in the game.

Yes gamification, the incorporation of game elements into an activity that people do not typcially consider a game\cite{2013Oxford}.  Why would software developers writing serious software be interested in this?  Our first step was to challenge this question through a survey that queried both whether developers would be interested in gamifying their workspace and what aspects of that space would be most interesting to them.

@@

The rest of the paper is organized as follows:  Section 2 reviews the preliminary survey-based assessment, Section 3 Discusses the design considerations for developing the Blaze tool/game, Section 4 Discusses the design of the study and participant selection, Section 5 is Related work, Section 6 results from the pilot, Section 7 Threats to Validity, and Section 8 has our Conclusions.

@@

Developers were asked to rank awards for different actions by which they would find the most desirable.  Overall 68\% of developers ranked receiving an award for consistently performing quality practices as either 1st or 2nd place.  This supports a conclusion that focusing on awarding quality practices would motivate developers to do them consistently.
%EMH connection?
\begin{itemize}
\item 95\% of developers responding would try tools and practices suggested by an automated recommendation system particularly if they were provided in context with the current work-flow pattern the developer follows.
\item Responses indicate a preference for team competition over individual awards (badges) and awards for quality practices ranked highest.  So, the tool will incorporate team competition ideas and recognize quality practices.
\item Over 90\% of respondents approved of sharing data with their team members
\end{itemize}

@@

To evaluate developers using recommended tools in Visual Studio, we categorized them by their tool name.  In the pilot we were interested in developers learning about and using the Sando search tool.  Sando is part of the recommended developer tool suite for ABB and supports structured navigation using code search \cite{Shepherd2012Sando}.

@@

Using these guidelines we assigned points to structured navigation commands which received 1 point each while unstructured navigation commands received 0 points.    We gave additional emphasis to using the Sando search tool by assigning it 10 points.  In creating levels, following the guidance from Hamari \cite{wbsnipes:Hamari2011Framework}, we built levels following an exponential curve.  Users initially level up after a working day or two extending to achieving the next level after a week or even a month following an exponential curve.  The levels make the game motivating for beginners and challenging for more experienced developers.  We designed the points and levels such that above average developers can complete all levels during the study period.

@@

After a week's data collection marking the end of the first stage, Blaze started to automatically pop-up a window containing a button for a web page with information.  The page included a link to download Sando  and information about using that tool in a video.  The sub-page on navigation contained a page of tutorial on the built-in structured navigation commands in Visual Studio.  The About Blaze page contained details of the point scoring system and background on Blaze's purpose.  The usefulness of the communication site is rated as part of our study completion survey.

@@

%\subsubsection{Observations About Developers` Habits}
%
%We are interested in understanding whether developers are consistently using practices that make them inefficient.  One way to check that is to determine whether our measurements are associated with specific developers in the study, in other words, can we independently classify the edits performed by each developer.  If this is possible, then we can discuss whether developer have consistent ways of navigating and understanding for each edit.  Thus the hypothesis is developers have consistent ways of navigating and understanding code while they are making changes.
%
%To test this hypothesis we construct a classification tree using Weka \cite{Hall2009WEKA} with the J48 classifier.  Slicing the data and training the model using 10-fold cross-validation, the model identifies 53\% of the 647 edit sessions with the correct developer among 6 participants.  The tree uses nearly all available fields including time for all categories and time of day for edits in its attempt to associate sessions with their developer generating a Kappa statistic of 0.403  and square root of the relative squared error at 101\%.  
%
%While there is some consistency to how developers work, we cannot say that individual developers are always consistent in their work patterns.  Thus developers are somewhat flexible in the commands they use when navigating and making changes to code and may be more open to adopting new tools and techniques.\

@@

Using Weka\cite{Hall2009WEKA}, we performed an attribute selection process to identify the most significant attributes related to edit session duration.  The attribute search method we chose was Greedy Step-wise search that selects attributes based on their correlation with the target and stops adding attributes when the next attribute decreases the correlation.    The process identified the top 5 out of 15 possible categories as follows:
\begin{enumerate}
\item Unstructured Navigation
\item Debug - events from running the debugger
\item Edit - events that modify the code
\item Other Actions - other events in visual studio that we do not sub-classify
\item Build - events from running a build
\end{enumerate}

Using R\cite{Rcitation}, we constructed a linear regression model for Unstructured Navigation to determine the correlation with edit duration.  The linear model found a positive correlation with unstructured navigation with a p-value <0.001 on 1694 degrees of freedom.  The model had an Adjusted R-squared of 0.64 showing a good portion of the variation in edit duration is explained by that category.

This analysis shows that we cannot disprove that unstructured navigation is correlated with edit duration, thus it is an important factor in developer productivity.  Combining unstructured navigation with reading describes a larger portion of the overall edit duration and motivates us towards making tools that improve both navigation methods and code comprehension.

\subsubsection{Observations on Navigation Ratio}

@@

Structured navigation events included: 
\begin{itemize}
\item Navigate To (Ctrl+,) is a fuzzy search interface that locates symbols matching your search terms.
\item Go To Definition (F12) brings up the code where the selected symbol is defined. 
\item View Call Hierarchy (Ctrl+K Ctrl+T) provides a two way analysis of a symbol's dependencies and uses. 
\item Class View (Ctrl+W, C) provides a browser and search function for classes and class hierarchy
\item Find All References provides a list of references for an object
\item Navigate to Event Handler brings up the event handler from the XAML editor
\item View Class Diagram generates a class diagram for the project
\item View Object Browser is a search tool for objects in the solution
\end{itemize}

@@

During the pilot we staged the deployment of information feedback to developers so they would first have access to learn about structured navigation and then get instant feedback in the form of points for using structured navigation commands.  Our hypothesis was that this instant feedback and information would result in an increase in the number of structured navigation commands used in Visual Studio.

To evaluate whether developers use more structured navigation when they receive instant feedback, we monitor the change in navigation ratio over the pilot study period.   Expecting that as developers learn about navigation commands and tools that count towards the score in the second week they will begin using the practices more.  In the third week we turn on points feedback and the leaderboard expecting this will encourage developers to use more structured navigation.

\begin{figure}
	\includegraphics[width=3.3in]{navratioboxplot_ann.pdf}
	\caption{Navigation Ratio by Week}
	\label{fig:navigationaverage}
\end{figure}

What we observed visually in Figure \ref{fig:navigationaverage} is that the structured navigation practice did not increase significantly from the beginning to the end of the study.  We evaluated the week over week differences using a Wilcoxon Rank sum test that compares week one to week two and week one to week three.    In week two when the developers were able to learn about structured navigation commands there was no significant change.  Also in week two, two developers stopped participating, one was working in another environment and the other dropped of for the rest of the study.  Week three there was a slight increase, however the increase is not significant.  Weeks four and five that follow the formal period of study show the navigation ratio continues in a similar range.  Overall, we did not see a significant effect towards increasing navigation ratio for the interventions Blaze provided to the environment.

\subsubsection{Observations on Points}

The developers using Blaze received feedback in the form of points given for using structured navigation commands and using the Sando search tool\cite{Shepherd2012Sando}.  Per the pilot design, the points display was disabled until the third week following installation so that a baseline without instant feedback could be established.  The points display and the navigation arrow were enabled in the third week to test whether the instant feedback was successful in driving increased use of the targeted features.  Observing the result shows that during the first two weeks, we see in Figure \ref{fig:pointsbyweek} developers were fairly consistent in their use patterns.  With the introduction of the points feedback, two developers` points accumulation spiked in week three due mostly to increased use of Sando.   We informed developers at the beginning of the pilot that Sando usage would receive 10 points and using built-in structured navigation commands would receive 1 point.

\begin{figure}
	\includegraphics[width=3.3in]{pointsbyweek_ann.pdf}
	\caption{Average Points per Day by Week}
	\label{fig:pointsbyweek}
\end{figure}

@@

@@

Following our data analysis, we conducted one on one feedback sessions with the pilot participants. Before initiating the discussion, we assured the individual developer that the data shared with them is confidential and we will not share the same with management. The participants received the feedback in the form of quantitative charts and written analysis of the information.  The key quantitative chart shown in Figure \ref{fig:developercomparison} compared their distribution of time in each category to the median for the group.

We discovered during these feedback sessions that developers did not spend the entire time in their desktop environment because they had a virtual machine environment available for testing. Thus categories like testing, debug, and reading were affected by whether or not they used the virtual environment or their desktop for these activities.  Individual differences in those categories are considered irrelevant and not discussed here. Because our game points analysis focused on navigation commands excluding reading, the presence of the virtual environment may explain how the data have missing values on some days.

We also discussed with one developer who`s build effort and the underlying build frequency was much higher than median as shown in Figure \ref{fig:developercomparison}.  We asked the developer to explain how building more frequently helps them. The developer mentioned that by doing more frequent build, (s)he is trying to ensure the correctness of the work completed so far. The developer feels that if (s)he resolves the bugs then and there, (s)he can do it faster as the code and logic is fresh in her memory. If they build less frequently, even though it reduces the build effort, there are more bugs each time and the developer often spends more time in identifying the problem area than fixing it. Hence the developer feels that more frequent builds improves their productivity.

In another case, we observed a higher level of unstructured navigation than median and asked the developer about that.  They explained that even though they agree that structured navigation is better than unstructured navigation, (s)he prefers unstructured navigation in some specific instances. The developer is the author of the code and (s)he knows the code completely. ``I know that the line I need to fix is in XYZ method and it is 4 blocks down from where I am currently. It takes much less time for me to reach the fix location using unstructured navigation. If the code is not written by me, I will prefer structured navigation.'' This instance cautions us to gather information on aspects such as code familiarity and developer experience with the product when assessing the results of a navigation study.

%Though commonly held opinion of Microsoft`s Clippy experience suggests this more invasive form of feedback is not always welcome, we shall list as future work providing more immediate suggestions on how developers can improve their navigation practices. 
Developers also reiterated that they wanted a different feedback experience from Blaze  such as immediate prompts to suggest navigation commands within the context of what they are working on.  Moving towards more immediate feedback, we shared a set of achievement badges tied to specific navigation commands  as we gave the more quantitative feedback to the developers.   Developers said that badges would motivate them to use some specific tools or commands more than points alone.   This contradicts a result from the pre-study survey where respondents rated badges as less important than team or individual competition in motivating them to improve.  The important factor in this case is badges serve as both a feedback mechanism and information showing which commands are important to use for improving structured navigation and their score.

@@

We conducted an in-person post pilot survey in order to cross check the quantitative data from Blaze with developers` perception of their practices and their demographic information.

The pilot survey queried developers` relevant demographic data to help understand some differences in results.  We asked developers to provide their years of experience so that we could evaluate any effects experience might have.  The developers in the study had between 7 and 14 years experience.  One developer who dropped out of the study said that their lower level of experience compared to their peers influenced their decision to stop participating.

We asked developers to rate their knowledge of the code they maintain.  Four of the six developers reported that they maintain code that they wrote while the two most experienced developers maintain code that they recently learned.  One developer reported during our individual analysis feedback that working in code they know well affects their navigation practices leading them to use more unstructured navigation.

On the assessment of structured navigation knowledge, half of the participants replied that they had very good prior knowledge about structured navigation and they use structured navigation as much as possible.  Other half of the participants have some understanding on structured navigation use it some of the time. Even developers whose navigation ratio  was lower than average, used structured navigation on occasion  as much as their colleagues.  While analyzing the points they scored with Blaze tool, we observed that the participants who rated their prior knowledge of structured navigation highly also scored higher in the pilot.

\begin{figure}
	\includegraphics[width=3.4in]{ElementInfluenceChart.pdf}
	\caption{Influence of Feedback Elements}
	\label{fig:elementInfluence}
\end{figure}

@@
